---
title: "Justin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
merged_df <- readRDS("data/merged_df.rds")
library(dplyr)
library(plotly)
library(maps)
library(mapdata)
library(mapproj)
library(ggmap)
library(ggplot2)
```

```{r box_data, echo = FALSE}

## Avg by_agi_range to show mean AGI and graduation rates by zip, including CORE region
agi_grad_df <- merged_df %>%
  filter(is.na(agi_range) & !is.na(zip_code)) %>%
  mutate(avg_agi = (agi_a * 1000) / return_c) %>%
  group_by(zip_code) %>%
  summarise(mean_agi = mean(avg_agi),
            mean_grad = mean(Graduation))

core_zip<- merged_df %>%
  filter(is.na(agi_range) & !is.na(zip_code)) %>%
  select(zip_code, CORE_region, system_name)

agi_grad_plot <- agi_grad_df %>%
  left_join(core_zip[c("zip_code", "CORE_region", "system_name")], by = "zip_code") %>%
  filter(!is.na(CORE_region))
```
 
## Graduation Rates by Region 


The boxplot below shows that, while most CORE regions in TN have a graduation rate around **90%**, we have some outliers in the Northwest and South Central regions.

The outlier in the Northwest region is *Humboldt City*, with a graduation rate of **68.5%**. This school system stands out in other troubling ways, which will be visited again later.

```{r boxplot, echo=FALSE, warning=FALSE}

## Boxplot
x_grad <- list(title = "", showticklabels = FALSE)
y_grad <- list(title = "Graduation Rate")

grad_box_plot<- plot_ly(agi_grad_plot, y = (~mean_grad), x = ~CORE_region, color = ~CORE_region,
                        type = "scatter",
                        mode = "markers",
                        alpha = 0.3,
                        text = ~paste('Zip Code: ', zip_code,
                                      '<br> System Name: ', system_name)) %>%
  layout(xaxis = x_grad, yaxis = y_grad, legend = list(orientation = 'h')) %>%
  add_trace(agi_grad_plot, y = ~mean_grad, color = ~CORE_region, type = "box")

grad_box_plot
```
As expected, a higher income from a particular county *does* result in a lower percentage of students that qualify as **"Economically Disadvantaged"** - students that qualified for free or reduced lunches.
```{r echo = FALSE, warning = FALSE}
# Cloropleth

## Total across zip code - removes agi ranges
zip_total <- merged_df %>% 
  filter(zip_code != 0 & is.na(agi_range))

## Filtering and summarising subject proficiency
chloro_df <- zip_total %>% 
  filter(is.na(agi_range)) %>% 
  select(CORE_region, system_name, county, Pct_ED, AlgI, AlgII, Math, 
         BioI, Chemistry, Science, ACT_Composite) %>%
  group_by(county) %>% 
  summarise_all(funs(mean)) %>% 
  filter(!is.na(county))

## Filter data to just include TN

TN_data <- map_data("state") %>% 
  filter(region =='tennessee')

TN_counties <- map_data("county") %>% 
  subset(., region == "tennessee")

TN_counties$subregion <- TN_counties$subregion %>% 
  gsub('de kalb', 'dekalb', .)

chloro_df$county_l <- sapply(chloro_df$county, tolower) %>% 
  gsub(' county','', .)

## Left join to merge polygon and county data

chloro_df <- left_join(x = TN_counties, y = chloro_df, by = c("subregion" = "county_l"))

## Switching to AGI data

agi_chloro_df <- zip_total %>% 
  filter(is.na(agi_range)) %>% 
  mutate(avg_agi = (agi_a * 1000)/return_c) %>% 
  select(CORE_region, system_name, county, avg_agi) %>%
  group_by(county) %>% 
  summarise_all(funs(mean)) %>% 
  filter(!is.na(county))

## Filter data to just include TN

agi_chloro_df$county_l <- sapply(agi_chloro_df$county, tolower) %>% 
  gsub(' county','', .)

## Left join to merge polygon and county data

agi_chloro_df <- left_join(x = TN_counties, y = agi_chloro_df, by = c("subregion" = "county_l"))

## Create maps

TN_agi_map <- ggplot() + 
  geom_polygon(data = agi_chloro_df, 
               aes(x = long, y = lat, group = group, fill = avg_agi),
               color = "white", size = 0.25) +
  coord_map() +
  scale_fill_distiller(name = "Average AGI", palette = "YlGn") +
  theme_nothing(legend = TRUE) +
  labs(title = "Average AGI") +
  theme(plot.title = element_text(hjust = 0.5)) 

TN_ed_map <- ggplot() + 
  geom_polygon(data = chloro_df, 
               aes(x = long, y = lat, group = group, fill = Pct_ED),
               color = "white", size = 0.25) +
  coord_map() +
  scale_fill_distiller(name = "Percent ED", palette = "YlGn", trans = "reverse") +
  theme_nothing(legend = TRUE) +
  labs(title = "% of Economically Disadvantaged Students") +
  theme(plot.title = element_text(hjust = 0.5))

TN_agi_map
TN_ed_map
```
 
## Impact of Economic Disadvantage 

In the plot below, we can see that the higher percentage of a student body that qualify for free or reduced lunches score lower in both Math and Science. 

*Humboldt City* school system had **97.9%** of their students qualify for free or reduced lunches. Students in the Humboldt City school system had a **43.8%** and **38.4%** proficiency rate in Science and Math, respectively. As you might recall, *Humboldt City* also had one of the lowest graduation rates in TN, at **68.5%**.

On the opposite side, *Lakeland City* reported a **14.7%** rate of Economically Disadvantaged students, while scoring **90.9%** and **88.2%** in Science and Math.


```{r, echo = FALSE}

## Total across zip code - removes agi ranges
zip_total <- merged_df %>% 
  filter(zip_code != 0 & is.na(agi_range))

## Filtering and summarising subject proficiency
subjects_df <- zip_total %>% 
  select(CORE_region, system_name, Pct_ED, AlgI, AlgII, Math, 
         BioI, Chemistry, Science) %>%
  group_by(CORE_region,
           system_name) %>% 
  summarise_all(funs(mean)) %>% 
  filter(!is.na(CORE_region))


## Melting data into tidy
ed_df <- reshape2::melt(subjects_df, id = c('CORE_region', 'system_name', 'Pct_ED'))
names(ed_df) <- c("CORE_region", "system_name", "Pct_ED", "Subject", "Pct_proficient")

ed_df <- ed_df %>% 
          filter(Pct_proficient > 0, system_name != "Achievement School District",
                 Subject == "Science" | Subject == "Math")

x_prof <- list(title = "% Economically Disadvantaged")
y_prof <- list(title = "Proficiency Rate")

prof_ed_plot<- plot_ly(ed_df, y = (~Pct_proficient), x = ~Pct_ED, color = ~Subject,
                        type = "scatter",
                        mode = "markers",
                        alpha = 0.7,
                        text = ~paste('CORE Region: ', CORE_region,
                                      '<br> School System: ', system_name,
                                      '<br> Subject: ', Subject)) %>% 
  layout(xaxis = x_prof, yaxis = y_prof, legend = list(orientation = 'v'))

prof_ed_plot
```